---
title: "RWorksheet_Songaling#4c"
author: "Charles Daniel Songaling"
date: "2023-11-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. Use the dataset mpg

#a. Show your solutions on how to import a csv file into the environment.

```{r}
library(readr)

mpgdata <- read.csv("mpg.csv")
mpgdata
```
#b. Which variables from mpg dataset are categorical?
```{r}
str(mpgdata)
```
#c. Which are continuous variables?
```{r}
continous_var <- c("disply","year","cyl","cty","hwy")
continous_var
```
#2. Which manufacturer has the most models in this data set? Which model has the most variations?Show your answer.
```{r}
library(ggplot2)

str(mpg)

manufacturer_model <- table(mpg$manufacturer)
manufacturer_max <- names(which.max(manufacturer_model))

cat("The manufacturer with the most models is:",manufacturer_max, "\n")

#Finding model with the most variations.
manufacturer_maxdata <- subset(mpg, manufacturer==manufacturer_max)
models <- table(manufacturer_maxdata$model)

Models_mostvariation <- names(which.max(models))
cat("The model with the most variations:", Models_mostvariation, "\n")
```
#a. Group the manufacturers and find the unique models. Show your codes and result.
```{r}
library(ggplot2)
library(dplyr)

manu_group <- mpg%>%
  group_by(manufacturer)%>%
  summarize(uniquemodel = n_distinct(model))

print(manu_group)
```
#b. Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r}
plot(manu_group$manufacturer, manu_group$uniquemodel,
     main = "Number of Unique Models by Manufacturer",
     xlab = "Manufacturer",
     ylab = "Number of Unique Models",
     col = "orange",
     names.arg = manu_group$manufacturer,
     las = 2, # Rotate the x-axis labels for better readability
     cex.names = 0.7)
     
#ggplot
library(ggplot2)
ggplot(manu_group, aes(x=manufacturer,y = uniquemodel))+
  geom_bar(stat = "Identity",fill = "purple")+
  labs(title = "Unique models per manufacturer",
       x = "Manufacturer",
       y = "Number of unique models")+
  theme_gray()
```
#2. Same dataset will be used. You are going to show the relationship of the model and the manufacturer.
#a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
```{r}
ggplot(mpg, aes(model, manufacturer)) + geom_point()

#It shows a scatter plot of the points where the model and manufacturer can be seen.
```
#b. For you, is it useful? If not, how could you modify the data to make it more informative?
```{r}
#For me, it is useful because it displays information about the models and their manufacturers.
```
#3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and its results.
```{r}
library(ggplot2)

top_20obs <- head(mpg,20)

ggplot(top_20obs, aes(x = model, y = year))+
  geom_point()+
  labs(title = "Model vs Year(Top 20 observations)",
       x = "Model",
       y= "Year")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))
```
#4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result.
```{r}
library(dplyr)
groups_modeldata <- mpg%>%
  group_by(manufacturer)
 

print(groups_modeldata)
```
#a. Plot using geom_bar() using the top 20 observations only. The graphs shoudl have a title, labels and colors. Show code and results.
```{r}
library(dplyr)
library(ggplot2)

data(mpgdata)

top_20obs <- head(mpg,20)

ggplot(top_20obs, aes(x = model, fill = model)) +
  geom_bar() +
  labs(title = "Number of Cars per Model (Top 20 Observations)",
       x = "Model",
       y = "Number of Cars") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  scale_fill_manual(values = rainbow(nrow(top_20obs)))
```
#b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.
```{r}
library(dplyr)
library(ggplot2)

data(mpgdata)

top_20obs <- head(mpg,20)

ggplot(top_20obs, aes(x = model, fill = model)) +
  geom_bar() +
  labs(title = "Number of Cars per Model (Top 20 Observations)",
       x = "Number of Cars",
       y = "Model") +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_fill_manual(values = rainbow(nrow(top_20obs))) +
  coord_flip()
```
#5. Plot the relationship between cyl - number of cylinders and displ - engine displacement usinggeom_point with aesthetic color = engine displacement. Title should be “Relationship between No. of Cylinders and Engine Displacement”.
```{r}
library(ggplot2)

data(mpgdata)


ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point() +
  labs(title = "Relationship between Number of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement") +
  scale_color_continuous(name = "Engine Displacement") +
  theme_minimal()
```
#a. How would you describe its relationship? Show the codes and its result.
```{r}
cyl_displ_relationship <- cor(mpg$cyl , mpg$displ)

cat("Correlation Coefficient:", cyl_displ_relationship, "\n")

library(ggplot2)
ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement") +
  scale_color_continuous(name = "Engine Displacement") +
  theme_minimal()

#As the number of cylinders increases, the engine displacement also increases.
```
#6. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped it with a continuous variable you have identified in #1-c. What is its result? Why it produced such output?
```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = year)) +
  geom_point() +
  labs(title = "Relationship between Engine Displacement and Highway MPG",
       x = "Engine Displacement",
       y = "Highway MPG") +
  scale_color_gradient(low="blue",high="violet") +
  theme_minimal()
```
#6. Import the traffic.csv onto your R environment.
#a. How many numbers of observation does it have? What are the variables of the traffic dataset. Show your answer.
```{r}
traffic_data <- read.csv("traffic.csv")

obsv_num <- nrow(traffic_data)
traffic_vars <- ncol(traffic_data)
var_names <- names(traffic_data)

cat("Number of Observations:", obsv_num, "\n")
cat("Number of Variables:", traffic_vars, "\n")
cat("Variables:", paste(var_names, collapse = ", "), "\n")
```
#b. subset the traffic dataset into junctions. What is the R codes and its output?
```{r}
traffic_junc <- c(1,2,3,4)

junction_subset <- traffic_data[traffic_data$Junction %in% traffic_junc, ]

junction_subset
```
#c. Plot each junction in a using geom_line(). Show your solution and output.
```{r}
library(ggplot2)

ggplot(traffic_data, aes(x = DateTime, y = Vehicles, color = as.factor(Junction))) +
  geom_line() +
  scale_color_manual(values = c("violet", "blue", "brown", "red")) + 
  labs(title = "Traffic by Junction",
       x = "Date Time",
       y = "Number of Vehicles") +
  theme_minimal() +
  facet_wrap(~Junction, scales = "free_y", ncol = 1) +
  guides(color = guide_legend(title = "Junction"))
```
#7. From alexa_file.xlsx, import it to your environment
#a. How many observations does alexa_file has? What about the number of columns? Show your solution and answer.
```{r}
library(readxl)

alexafile <- read_excel("alexa_file.xlsx")

#Observations and columns
alexaobs <- nrow(alexafile)

alexa_cols_obs <- ncol(alexafile)

cat("The number of observations on alexa_file is:",alexaobs,"\n")
cat("The number of coloumns on alexa_file is:",alexa_cols_obs,"\n")
```
#b. group the variations and get the total of each variations. Use dplyr package. Show solution and answer.
```{r}
library(dplyr)

variations_group <- alexafile%>%
  group_by(variation)%>%
  summarise(totalcount_ = n())
variations_group
```
#c. Plot the variations using the ggplot() function. What did you observe? Complete the details of the graph. Show solution and answer.
```{r}
library(ggplot2)

ggplot(alexafile, aes(x = variation)) +
  geom_bar() +
  labs(title = "Distribution of Variations",
       x = "Variation",
       y = "Count") +
  theme_minimal()
```
#d. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the graphs. Show your answer and solution.
```{r}
alexafile$date <- as.Date(alexafile$date)

alexafile$month <- format(alexafile$date, "%m")

months <- alexafile %>%
  group_by(month) %>%
  summarise(num_reviews = n())


months_review <- table(months)

ggplot(months, aes(x = month, y = num_reviews, group = 1)) +
  geom_line(color = "green") +
  labs(title = "Number of Verified Reviews Per Month",
       x = "Month of 2017", y = "Number of Reviews")
```
#e. Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a graph to show its relationship. Show your solution and answer.
```{r}
library(dplyr)

ggplot(alexafile, aes(x = variation, y = rating, fill = variation)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  scale_fill_manual(values = rainbow(n = length(unique(alexafile$variation)))) +  
  labs(title = "Relationship Between Variations and Ratings",
       x = "Variations",
       y = "Average Rating") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ratings <- alexafile %>%
  group_by(variation)%>%
  summarise(avearage_rating = mean(rating, na.rm = TRUE))


high_rate <- ratings %>%
 filter(average_rating  == max(average_rating)) %>%
  pull(variation)

cat("The Variations that got the highest in rating is:",high_rate)
```