---
title: "RWorksheet_Songaling#4c"
author: "Charles Daniel Songaling"
date: "2023-11-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. Use the dataset mpg

#a. Show your solutions on how to import a csv file into the environment.

```{r}
library(readr)

mpgdata <- read.csv("mpg.csv")
mpgdata
```
#b. Which variables from mpg dataset are categorical?
```{r}
str(mpgdata)
```
#c. Which are continuous variables?
```{r}
continous_var <- c("disply","year","cyl","cty","hay")
continous_var
```
#2. Which manufacturer has the most models in this data set? Which model has the most variations?Show your answer.
```{r}
library(ggplot2)

str(mpg)

manufacturer_model <- table(mpg$manufacturer)
manufacturer_max <- names(which.max(manufacturer_model))

cat("The manufacturer with the most models is:",manufacturer_max, "\n")
```
#a. Group the manufacturers and find the unique models. Show your codes and result.
```{r}

```
#b. Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r}
barplot(manufacturer_model$manufacturer, manufacturer_model$uniquemodel,
     main = "Unique models per manufacturer",
     xlab = "Manufacturer",
     ylab = "Number of unique models")
     
```
#2. Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.
```{r}

```
#a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
```{r}

```
#b. For you, is it useful? If not, how could you modify the data to make it more informative?
```{r}

```
#3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and its results.
```{r}
library(ggplot2)

top20 <- head(mpg,20)

ggplot(top20, )

```
#4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result.
```{r}

```
#a. Plot using geom_bar() using the top 20 observations only. The graphs shoudl have a title, labels and colors. Show code and results.
```{r}

```
#b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.
```{r}

```
#5. Plot the relationship between cyl - number of cylinders and displ - engine displacement usinggeom_point with aesthetic color = engine displacement. Title should be “Relationship between No. of Cylinders and Engine Displacement”.

#a. How would you describe its relationship? Show the codes and its result.
```{r}

```
#6. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped it with a continuous variable you have identified in #1-c. What is its result? Why it produced such output?
```{r}

```
#6. Import the traffic.csv onto your R environment.
#a. How many numbers of observation does it have? What are the variables of the traffic dataset the Show your answer.
```{r}

```
#b. subset the traffic dataset into junctions. What is the R codes and its output?
```{r}

```
#c. Plot each junction in a using geom_line(). Show your solution and output.
```{r}

```
#7. From alexa_file.xlsx, import it to your environment
#a. How many observations does alexa_file has? What about the number of columns? Show your solution and answer.
```{r}

```
#b. group the variations and get the total of each variations. Use dplyr package. Show solution and answer.
```{r}

```
#c. Plot the variations using the ggplot() function. What did you observe? Complete the details of the graph. Show solution and answer.
```{r}

```
#d. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the graphs. Show your answer and solution.
```{r}

```
#e. Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a graph to show its relationship. Show your solution and answer.
```{r}

```


