---
title: "RWorksheet_Songaling#6"
author: "Charles Daniel Songaling"
date: "2023-12-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.Create a data frame for the table below. Show your solution.
```{r}
Student_Scores <- data.frame(
  Student = c(1:10),
  Pre_test = c(55,54,47,57,51,61,57,54,63,58),
  Post_test = c(61,60,56,63,56,63,59,56,62,61)
)
Student_Scores

names(Student_Scores) <- c("Student","Pre-test","Post-test")
```
#a. Compute the descriptive statistics using different packages (Hmisc and pastecs). Write the codes and its result.
```{r}
library(Hmisc)
library(pastecs)

#Hmisc
describe(Student_Scores)


#pastecs
stat.desc(Student_Scores)
```

2. The Department of Agriculture was studying the effects of several levels of a fertilizer on the growth of a plant. For some analyses, it might be useful to convert the fertilizer levels to an ordered factor.
The data were 10,10,10, 20,20,50,10,20,10,50,20,50,20,10.

#a. Write the codes and describe the result.
```{r}
Fertilizer_levels <- c(10,10,10, 20,20,50,10,20,10,50,20,50,20,10)

ordered_factors <- ordered(Fertilizer_levels, levels = c(10,20,50))

ordered_factors

#The numbers inside the square bracket represent the observations or data points and below are the levels. The arrow is the one that indicates the order from 10 less than to 20 and 20 less than to 50.
```

3. Abdul Hassan, president of Floor Coverings Unlimited, has asked you to study the exercise levels undertaken by 10 subjects were“l”, “n”, “n”, “i”, “l” , “l”, “n”, “n”, “i”, “l” ; n=none, l=light, i=intense.

#a. What is the best way to represent this in R?
```{r}
Exercise_levels <- c("l", "n", "n", "i", "l", "l", "n", "n", "i", "l")


Exercisefactor <- factor(Exercise_levels, levels = c("n", "l", "i"))

Exercisefactor
```

4. Sample of 30 tax accountants from all the states and territories of Australia and their individual state of origin is specified by a character vector of state mnemonics as:

#a. Apply the factor function and factor level. Describe the results.
```{r}
State <- c("tas", "sa", "qld", "nsw", "nsw", "nt", "wa", "wa", "qld",
"vic", "nsw", "vic", "qld", "qld", "sa", "tas", "sa", "nt",
"wa", "vic", "qld", "nsw", "nsw", "wa", "sa", "act", "nsw",
"vic", "vic", "act")

FactorState <- factor(State, levels = c("act", "nsw", "nt", "qld", "sa", "tas", "vic", "wa"))

FactorState

#The numbers inside the square brackets are the observations and below it are the levels. The levels represent the specified geographic regions of Australia.
```

5. From #4 - continuation:
Suppose we have the incomes of the same tax accountants in another vector(insuitably large units of money)

#a. Calculate the sample mean income for each state we can now use the special function tapply():
```{r}
incomes <- c(60, 49, 40, 61, 64, 60, 59, 54,
62, 69, 70, 42, 56, 61, 61, 61, 58, 51, 48,
65, 49, 49, 41, 48, 52, 46, 59, 46, 58, 43)

incmeans <- tapply(incomes, FactorState, mean)
incmeans


#b.  Copy the results and interpret.
#We can see that it calculates the means of every states.
```

6. Calculate the standard errors of the state income means (refer again to number 3).

#a. What is the standard error? Write the codes.
```{r}
stdError <- function(x) sqrt(var(x)/length(x))
incster <- tapply(incomes, FactorState, stdError)
incster

#b. Interpret the result.
#In No.5, we see the means of every states while in No.6, we calculate the standard error of each states. The standard errors provide a measure of the uncertainty associated with the sample mean incomes for each state. Lower standard errors indicate more precise estimates, while higher standard errors suggest greater variability in the estimates.
```

7. Use the titanic dataset.

#a. subset the titatic dataset of those who survived and not survived. Show the codes and its result.
```{r}
library(titanic)

data("titanic_train")

Survived <- subset(titanic_train, Survived == 1)

Did_not_Survive <- subset(titanic_train, Survived == 0)

head(Survived)
head(Did_not_Survive)
```

8. The data sets are about the breast cancer Wisconsin. The samples arrive periodically as Dr. Wolberg reports his clinical cases.

#a. describe what is the dataset all about.
```{r}
breastcancer_data <- read.csv("breastcancer_wisconsin.csv")

str(breastcancer_data)
head(breastcancer_data)
summary(breastcancer_data)

#The dataset is all about data regarding breast cancer.
```
#d. Compute the descriptive statistics using different packages. Find the values of: d.1 Standard error of the mean for clump thickness. d.2 Coefficient of variability for Marginal Adhesion. d.3 Number of null values of Bare Nuclei. d.4 Mean and standard deviation for Bland Chromatin d.5 Confidence interval of the mean for Uniformity of Cell Shape
#d. How many attributes?
```{r}
library(psych)

clump_thickness <- breastcancer_data$ClumpThickness
marginal_adhesion <- breastcancer_data$MarginalAdhesion
bare_nuclei <- breastcancer_data$BareNuclei
bland_chromatin <- breastcancer_data$BlandChromatin
uniformity_cell_shape <- breastcancer_data$UniformityCellShape

#d.1 Standard error of the mean for clump thickness.

SE_clumpthickness <- sd(clump_thickness) / sqrt(length(clump_thickness))
SE_clumpthickness

#d.2 Coefficient of variability for Marginal Adhesion.
CV_marginaladhesion <- sd(marginal_adhesion) / mean(marginal_adhesion)
CV_marginaladhesion

#d.3 Number of null values of Bare Nuclei.
nullval_barenuclei <- sum(is.na(bare_nuclei))
nullval_barenuclei

#d.4 Mean and standard deviation for Bland Chromatin
mean_blandchromatin <- mean(breastcancer_data$bland_chromatin)
sd_blandchromatin <- sd(breastcancer_data$bland_chromatin)
mean_blandchromatin
sd_blandchromatin

#d.5 Confidence interval of the mean for Uniformity of Cell Shape
ci_uniformitycellshape <- tryCatch(
  t.test(breastcancer_data$`uniformity_cell_shape`)$conf.int,
  error = function(e) NULL
)
ci_uniformitycellshape

#d. How many attributes?
num_attributes <- ncol(breastcancer_data)
print(paste("Number of Attributes:", num_attributes))
```
#e. Find the percentage of respondents who are malignant. Interpret the results.
```{r}
percentage_malignant <- (sum(breastcancer_data$Class == "Malignant") / nrow(breastcancer_data)) * 100
print(paste("Percentage of Respondents who are Malignant:", percentage_malignant))
#There are no respondents who are malignant.
```

9. Export the data abalone to the Microsoft excel file. Copy the codes.
```{r, error=TRUE}
library(AppliedPredictiveModeling)

data("abalone")


library(openxlsx)

write.xlsx(abalone, file = "abalone.xlsx")

View(abalone)
head(abalone)
summary(abalone)
```